---
title: "DRY coding"
author: "Data Science in Biomedicine - Week 5"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```
## Introduction

We'll be using an in-built data set from the MASS package. But, because the MASS package can be a bit problematic to install, I've provided the data for you. It's in .rds format, so we load it in using the `read_rds()` function.

For more info about the dataset, head to https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/survey.html

```{r}
library(tidyverse)
library(knitr)

survey <- read_rds("./data/survey.rds")
```

## Exercise 1
Explore the data - what variables are in there, what variable types, how many variables and observations?
```{r}
str(survey)
```


## Exercise 2
Round every numeric variable to 0 decimal places.

```{r}
survey_rounded <- survey %>% 
  mutate(across(where(is.numeric), round, 0))
```


## Exercise 3
Using the rounded dataframe you created in Exercise 2, calculate the mean of each numeric variable in the dataset. Use kable() to insert a nicely formatted table into your R Markdown document.

```{r}
mean_survey <- survey_rounded %>% 
  select(where(is.numeric)) %>% 
  map_df(~mean(.x, na.rm = TRUE)) %>% 
  kable(
    col.names = c("Wr.Hnd","Survived", "Died", "Total", "Percent Survived"),
    caption = "Mean of Numeric Variables",
    align = "ccccc"
  )

mean_survey
```


## Exercise 4
Now make exactly the same table, using the pivot functions. Again, use kable() to insert a nicely formatted table into your R Markdown document.
```{r}
mean_survey_2 <- survey_rounded %>% 
  select(where(is.numeric)) %>% 
  pivot_longer(
    cols = everything()) %>% 
  group_by(name) %>% 
  summarise(Mean = mean(value, na.rm = TRUE)) %>%
  pivot_wider(
    names_from = name,
    values_from = Mean
  ) %>% 
  kable(
    col.names = c("Wr.Hnd","S", "D", "T", "P"),
    caption = "Mean of Numeric Variables",
    align = "ccccc"
  )
  mean_survey_2
```

## Exercise 5
Students had initially expressed their height in either imperial or metric units, but someone converted everything into centimeters. Create a new variable, Height_original, which will show height in centimeters for students who expressed their height in metric units, and height in inches for students who prefer imperial units. 

NOTE: This one is a bit of a challenge. You may want to check out the `case_when()` function.

```{r}
survey_utate <- survey %>% 
  mutate(Height_original = case_when(M.I == "Imperial" ~ Height*0.3937,
                                     M.I == "Metric" ~ Height))
```


## Exercise 6
Create a table that shows the mean and standard deviation of height - in centimeters for students who used metric units, and in inches for students who used imperial units. Use kable() to insert a nicely formatted table into your R Markdown document.

```{r}
survey_utate_2 <- survey_utate %>% 
  select(M.I, Height_original) %>% 
  group_by(M.I) %>% 
  summarise(Mean = mean(Height_original),
            Standard_Deviation = sd(Height_original)) %>% 
  filter(!is.na())
```


## Exercise 7
Create a contingency table to see the relationship between exercise and smoking. Pivot your table to make it easier to read, and then use kable() to insert a nicely formatted table into your R Markdown document.

```{r}

```


## Exercise 8
Create a grid of Pulse histograms - one histogram for each combination of Sex and Exercise.

```{r}

```


## Exercise 9
Create a new variable, Hnd.diff, that shows the difference in span between the writing and non-writing hand. Create a set of boxplots of this variable, with a separate boxplot for those who write with their right hand, and another one for those who write with their left hand.

```{r}

```

